---
title: "Generic Ion Leakage Plot"
author: "Marijn Knip"
date: "21 December 2017"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
rm(list=ls())
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

## Simple Line graph for Ion Leakage experiments

This script generates simple line graphs for data generated in an ion-leakage experiment using leaf discs.
This script reads *.csv files in long format located in the same folder as the script file.
In this section different elements in the output figure can be set.

```{r}
## install libraries below, if needed
## install.packages("ggplot2")
## install.packages("doBy")
## install.packages("rstudioapi")
library(ggplot2)
library(doBy)
library(rstudioapi)

##set working directory to location of this .R file and read datafile
##setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
  iondata <- read.table("testdata.csv", ##change to your *.csv
                     header = T,
                     sep = ";",
                     dec = "."
)

##insert plot title, axis titles, legend text etc. here
plottitle=    "Rx1-induced ion leakage"
plotsubtitle= "virulent vs. avirulent coat protein"
legendtitle=  "Constructs"
xaxislabel=   "Time after 20 \U03BCM dexamethasone treatment (hours)"
yaxislabel=   "Conductivity (\U03BCS/m)"
legendname1=  "DEX::CP105"
legendname2=  "DEX::CP106"
legendname3=  "Sample2"
legendname4=  "Sample2+treament"
##legendname works for 4 entries, change code at p1 to change number

##summarizing data
data <- summaryBy(Conductivity ~ Timepoint + Construct, data=iondata, FUN=c(length,mean,sd))

summary(data)

```

## Building the plot
In the section below is the code to generate the simple plot:

```{r}
##drawing plot > Mean values incl. stdev.
#line graph with dots, slightly transparent (alpha)
# set position dodge to move data pounts slightly
##Using scale_colour_manual (matching brewer "spectral" colour), as using only two colors brewer with spectral palette doesn't give red/blue 
pd <- position_dodge(0.00)
p1 <- ggplot(data, aes(x=Timepoint, y=Conductivity.mean,color= Construct)) + 
  geom_errorbar(aes(ymin=Conductivity.mean-Conductivity.sd, ymax=Conductivity.mean+Conductivity.sd), width=0.2, position=pd, alpha = 0.4) +
  geom_line(size=1, position = pd) +
  scale_colour_manual         (labels = c(legendname1,legendname2),
                                values = c("#2b83ba","#d7191c")) +
  geom_point(size=3, fill="white", alpha = 0.4, position = pd)
plot(p1)
```

## Plot design
Now the plot is generated we improve the design:

```{r}
#enhancing plot design
#setting labels and titles
p2 <- p1 + labs(x=xaxislabel, 
                y=yaxislabel,
                fill=legendtitle,
                title=plottitle,
                subtitle=plotsubtitle
               )
               
#changing Theme
p3 <- p2 +theme(
  plot.title = element_text(hjust = 0.5 , color="black", size=24, face="bold"),
  plot.subtitle = element_text(hjust = 0.5 , color="black", size=16, face="italic"),
  axis.title.x = element_text(color="black", size=14, face="plain"),
  axis.title.y = element_text(color="black", size=14, face="plain"),
  panel.background = element_rect(fill = "#f9f9f9",
                                  colour = "#f9f9f9",
                                  size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                  colour = "white"), 
  panel.grid.minor = element_blank(),
  text = element_text(size=14, face = "bold"), 
  legend.key = element_blank(),  ## Get rid of the legend box 
  legend.title = element_text(size = 12, face="bold"),
  legend.text = element_text(size = 10, face="italic"),
  legend.background = element_rect(fill=alpha(0.0001))
)
plot(p3)
```